### 2016-10-03 Paper_Final.tex
Updated to make a few final tweaks to writing.

\section*{Introduction}
Cabinet formation has long been considered from the perspective of parties and coalitions. When forming a cabinet, parties must ask who to include to please constituencies, geographies, factions within the party, and coalition partners. But parties are not monoliths, they are aggregations of members with individual preferences and goals. (M{\"u}ller and Str{\o}m 2000) Among the goals sought by any legislator are ability to achieve policy outcomes; personal and career satisfaction; service to party; and re-election. (Fenno 1978) Joining cabinet plausibly assists with the first three of these goals, but \textit{do cabinet ministers additionally receive an electoral return in the form of higher vote shares or increased electoral survival}? This question is a relevant consideration for MPs considering joining a cabinet, parties considering electoral strategy, and researchers seeking to understand the behaviour of legislators and the incentives they face.

Past evidence suggests that governing parties incur a ``cost of governing'', which individual members can mitigate or overcome by means of developing a personal vote. (Narud and Valen 2008; Akkerman and Lange 2012) This paper extends Martin (2016), who found a strong electoral effect associated with joining cabinet in Ireland. I test two national cases: Canada and New Zealand. The Canadian case reveals a robust but modest effect across specifications, while New Zealand demonstrates no consistent effect.

I continue by presenting relevant past work; explaining case selection, dataset construction, and research design; presenting empirical results; and conclude by discussing possible reasons for dissonance between the Canadian and New Zealand results.

@@

Downs (1957), Key (1964), and Mayhew (1974) are among early authors who explained legislative and party behaviour as being animated by desire for re-election. Mayhew articulates three mechanisms by which legislators earn re-election: advertising, credit claiming, and position taking. Fenno (1978) takes a broader view of legislator behaviour and expands on the ways in which legislators form relationships with their constituents to secure re-election. An essential tension exists: on one hand, incumbent governments pay an electoral cost for governing. (Mershon 1996) On the other hand, a widely documented incumbent advantage makes staying elected an easier proposition than winning an open election. Incumbents deter competition and enjoy resource advantages. (Erikson 1971; Carson 2005; Ban, Llaudet, and Snyder 2016; Gelman and King 1990)

Legislators seek to maximize their incumbency advantage while minimizing their costs of governing. Cain, Ferejohn, and Fiorina (1984) articulate how constituency service can be used to build a ``personal vote''. The incentives to earn a personal vote vary by electoral and institutional context. Systems where voters choose candidates directly offer the highest incentive for a personal vote, because members can be rewarded or protected by developing a reputation beyond that of their party. (Carey and Shugart 1995) The tools available to members also vary by context. Parliamentary systems impose whipped voting and use scheduling rules to prohibit backbenchers from advancing substantive business, theoretically limiting the ability of members to distinguish themselves outside the party. This has ramifications for the way in which legislators see their roles, what they hope to accomplish in a legislature, if they will continue to serve, and how they situate themselves with respect to their voters.

Studies of particularistic legislative spending find that legislators gain electoral advantage through the delivery of spending and targeted policy goods. In a US congressional context, studies focus on the ability of individual legislators to secure spending for their home districts through amendments to broader spending bills and through committee membership. Parliamentary systems limit the ability of backbenchers to achieve these benefits on their own. Instead, the theory proposes that a cartel logic takes hold. In one telling of the story, the governing party uses distribution of particularistic benefits to benefit all its members. But this effect does not vary from member to member and would be more meaningfully targeted at marginal ridings: it is a party-level effect. In the cabinet story, cabinet ministers have control over the shape of policy and spending within their particular ministry. Cabinet ministers support each other through the targeted provision of these goods in concert with one another: what Martin (2016) terms ``executive particularism.'' Preliminary work in a variety of national settings finds that such particularistic spending occurs at the cabinet level. (John, Ward, and Dowding 2004; Denemark 2000)

Martin's ``Policy, Office, and Votes: The Electoral Value of Ministerial Office'' is the only published work that focuses on quantifying electoral advantage conferred by cabinet membership. Martin analyzes Irish data beginning in 1980. The empirical component tests two simple models. A cross-sectional association finds that cabinet ministers on average earn 8.5\% more of their electoral quota\footnote{Martin's dependent variable is percentage of minimum electoral quota earned from first-choice ballots, a choice that is not relevant in the electoral contexts studied in this paper.} than like non-ministers. A model with change in vote share as the dependent variable finds that while government members face an 8\% cost of governing, cabinet ministers erase this cost with an 8.7\% return from cabinet membership. Martin ends with a claim that ``the broad theory... is likely to be unassailable'' and an invitation to extend the analysis to different institutional and electoral environments, including mixed-member proportional systems.

In addition to executive particularism, exposure may also be a driver of electoral success. Additional media exposure and the opportunity to hold credit claiming events in concert with spending or legislation may assist cabinet ministers. Evidence indicates that credit claiming is an important part of the ability to capitalize electorally on the delivery of spending or policy goods. (Grimmer, Messing, and Westwood 2012; Bickers, Evans, Stein, and Wrinkle 2007) To the extent that ministers enjoy a differential ability to claim credit, they should reap rewards for doing so.

\section*{Case Selection}
The ideal cases to test for a cabinet effect are Westminster parliamentary democracies with a long history of competitive elections, clearly defined cabinets drawn from elected lower houses, incumbents elected primarily from district-level elections, a norm against district-switching by incumbents, and a tendency towards single-party government. Selecting for the latter allows the research to avoid engaging the complexities of cabinet appointment in a coalition setting, intra-election coalition changes. (Laver and Shepsle 1996) I narrow my choice of cases to Australia, Canada, Ireland (studied by Martin), New Zealand, and the United Kingdom. I selected Canada and New Zealand from these initial cases.

Canada is a Westminster parliamentary democracy established in 1867. It features a 338 member lower house, an appointed upper house that as a matter of convention does not impede government business, and a first-past-the-post electoral system. Canadian has never formed a coalition government, preferring by convention minority governments. The dominant social cleavages are ethno-linguistic between English and French regions of Canada, followed by core-periphery inter-provincial tensions. Canada has historically been governed by one of two parties: the Liberal Party and the Progressive Conservative Party (after 2004, the Conservative Party). Minor regional parties have achieved some electoral success. The effective number of political parties\footnote{A measure of party competition, where $N = \frac{1}{\sum_{i=1}^n p_i^2}$, with parties 1 through \textit{n} each occupying a $p_i$ share of seats. The measure was developed by Laakso and Taagepara (1979).} during the time studied varies from 1.5 to 3.2. Once they have contested an election, candidates do not switch ridings in subsequent elections, although parachuting of star candidates into politically opportune ridings occurs among first-time candidates. (Koop and Bittner 2011) Elections in Canada are relatively low cost with bans on corporate and union donations, extensive public funding, and spending caps. Major party candidates typically spend in the order of \$75,000 per four-year election.

Canadian Prime Ministers enjoy broad latitude to make cabinet appointments, subject only to informal norms with respect to the representation of each province in cabinet, the allocation of certain portfolios to relevant areas (for example, the Fisheries ministry is traditionally given to an MP from one of Canada's coastal provinces), and increasingly representation of gender, immigrant groups, and occupational diversity. A survival analysis of cabinet entry in Canada found that gender, legal experience, education, age, and past ministerial tenure were major predictors of selection, while margin of victory and regional support were not. (Kerby 2009) Over time, cabinets in Canada have become increasingly professionalized and technocratic, a trend also observed in other developed democracies (Lammers and Nyomarkay 1982; Pekkanen, Nyblade, and Krauss 2013; Berlinski, Dewan, Dowding, and Subrahmanyam 2009)

New Zealand is a Westminster parliamentary democracy established in 1852. Historically it featured a lower house with fewer than 100 members, a perfunctory upper house abolished in 1950, and a first-past the post electoral system with reserved electorates for voters who self-identified as indigenous Maori. Dissatisfaction with the two major political parties, National and Labour, and repeated majority governments elected by weak pluralities of popular vote support, New Zealand enacted electoral reform beginning in the 1996 national election. The current electoral system is closed-list mixed-member proportional with a low minimum vote threshold for representation, and consists of approximately 121 seats including 71 electorate seats. The effective number of political parties varied from 1.76 to 2.16 before MMP and 2.78 to 3.76 after MMP. Governments post-MMP have either featured formal supply agreements or coalitions with minor parties. Elections in New Zealand are extremely low-expense, with major party candidates spending in the order of \$20,000 per three-year election.

In principle, Labour governments in New Zealand historically allowed their entire caucus to vote on cabinet membership, but in practice Prime Ministers are allowed considerable leeway to expand or shrink the size of cabinet and allocate portfolios among those selected. (Alley 1989; Wood 1989) Cabinets typically include representation from both of New Zealand's islands and from urban and rural areas. The Minister of Maori Affairs position is often filled by a Maori legislator, and increasingly the representation of women in cabinet has been a relevant factor. (Curtin 2015)

Both countries share a variety of cultural and institutional factors: both feature high levels of parliamentary discipline and practice cabinet collective responsibility. Expulsion from cabinet, caucus, and denial of renomination are the primary punishments available to enforce discipline. Neither country has a formal primary election mechanism, instead delegating candidate nominations to district-level parties, who by custom do not remove incumbents for reasons other than non-compliance with party leadership or scandal. (Levine 1979; Wood 1989) Canada has a norm of allowing ``conscience votes'' on social issues when such votes do not impede legislation from proceeding. Finally, both cases are known to practice ``government by cabinet'', with inert backbenches and powerful cabinets. (Henderson 1989)

New Zealand and Canada differ in terms of incumbency. Canada by some measures has the lowest level of incumbency among developed countries, while New Zealand has above-average incumbency. (Matland and Studlar 2004) Incumbency has a profound effect on cabinet selection: in the Canadian low-incumbency situation, ascendant governments are typically forced to assemble a cabinet while appointing less experienced members of caucus. (Kerby 2015)

\section*{Data}
The dataset I use is novel. Existing election data suffers from limitations that renders it inappropriate for use to study my questions of interest. Much of the data is aggregated above the constituency level; constituency level data is often missing names and characteristics of individual candidates necessary to establish incumbency across redistricting or district switching and convert the data into a panel; and data typically does not include cabinet membership covariates.

As a result, I assembled a dataset containing full, constituency-level electoral data for all candidates and legislators in Canada from 1867 forward, and New Zealand from 1945 forward.\footnote{Canadian cabinet ministers before 1931 were expected to immediately resign from the House of Commons upon being appointed and then run in by-elections to regain the seat they resigned. To avoid the complexity of this feature of the data and the tail end of a party system transition occurring in the 1930s, I discard data before the 1945 federal election.} Canadian data was obtained by writing a series of web scrapers and harvested from Canadian Library of Parliament sources. New Zealand data was assembled using previously published volumes\footnote{Wood 1996; Norton 1988}, paper returns supplied by the NZ Electoral Commission and digitized for the first time for this project, and web scrapers for data from 1996 onwards.

@@

This paper tests one major hypothesis: Do cabinet ministers enjoy an electoral advantage compared to backbenchers?

The primary limitation when considering this question is severe endogeneity. The ``gold standard'' to tackle this problem would be to experimentally appoint cabinet ministers through random assignment and measure causal effects. Reality is not so kind: Parties select cabinet ministers to satisfy objectives including rewarding key constituencies, demographics, and policy demanders within the party and coalition partners. Moreover, parties have access to private signals about incumbent quality learned during the recruitment, vetting, and campaigning process that are not fully accounted for at the time of their first election. These latent candidate qualities may become obvious during the process of legislating, but unrelated to cabinet appointment: Koop and Bittner (2011) argue that ``star candidates'' are much more likely to be appointed to cabinet. Given these concerns, the claims made here cannot be said to be causal: effects here are associations, suggestive of a relationship, but not decisive.

Tables in this paper present a main model and a model that ``replicates'' Martin's comparable design. I run alternative specifications to test robustness of the observed effects and include these in Appendices 1-4.

@@

George Box remarked that ``All models are wrong, but some are useful''. This aphorism is a reminder that model specification is not designed to uncover ``truth'', it is designed to uncover useful patterns in the data at hand to support or refute theory. In keeping with this observation, it is likely that the models presented here omit variables due to limitations of data, theory, or imagination. Of particular note, the scope of this project rendered it impossible to collect richer biographical covariate data on candidates. Subsequent research that can gather such data could explain residual variation in candidate quality and more clearly isolate the effect of interest or allow a more robust selection on observables causal identification strategy. In the mean time, the modeled variables are as follows:

\setstretch{1.5}
\begin{itemize}
\item \textit{VotePct}: The vote share of the candidate.
\item \textit{CabinetNow}: The candidate was a member of the cabinet during the preceding term. I exclude Ministers Outside Cabinet, Junior Ministers, Ministers Without Portfolio, and other minor quasi-cabinet roles but include Deputy Prime Ministers and Attorneys General.\footnote{I also do not model ministerial exit from cabinet or role-switching mid-term, so in principle some of the \textit{CabinetNow} incumbents have been removed from cabinet before the election. Evidence suggests that ministerial exit is predicted by poor performance (Berlinkski, Dewan, and Dowding 2010) or scandal, (Dewan and Myatt 2007; Dewan and Dowding 2005) so this modeling decision should cause effect estimates to be biased conservatively.}
\item \textit{CabinetImportant}: The candidate was an ``important'' minister of the cabinet during the preceding term. ``Importance'' is a subjective definition and includes Defence, Health, Justice, Finance, Public Works, Revenue, and Foreign ministries. This variable exists to capture heterogeneity in the cabinet effect; senior ministers are presumed to have more discretion with respect to particularistic spending, more control over policy outcomes, and greater capacity to cultivate a personal vote through media exposure and credit claiming.
\item \textit{CabinetPM}: Dummy variable for the Prime Minister, regardless of whether they were an incumbent member of the legislature. The inclusion of a control for Prime Ministers is intended to absorb some of the positive residual expected for a sitting Prime Minister and not properly attributed to a cabinet effect.
\item \textit{Incumbent}: Candidate is an incumbent member of the legislature. I treat candidates as incumbent even if redistricting occurs, or if the candidate was previously elected by party list and now contests an electorate. Incumbency is known to be a strong factor in probability of re-election.
\item \textit{TermsServed}: The total unbroken number of terms served by a candidate. Tenure does not reset if an incumbent resigns and runs in a by-election, but does reset if a candidate loses or resigns and later returns to the legislature. I treat by-elected members as serving a full time. Tenure is present in the model to capture the accumulated effects of developing a personal vote (or, at least, a reputation of competency) in an incumbent's district over time.
\item \textit{Age}: The age of the candidate at the time of the election. Only available for a subset of Canadian candidates who were elected to office at least once. Age was gathered to allow for a proxy for signals of candidate quality including occupation, past electoral and career experience.
\item \textit{Elected}: Binary indicator for election outcome.
\item \textit{Party}: Party label the member ran under. I recode minor parties or splinter faction labels to be Independent.
\item \textit{PartyInGovt}: Candidate is running for the party currently governing. In New Zealand post-1996, I include junior coalition partner parties but not parties in supply agreements with the governing party. This variable was constructed in order to account for the ``electoral costs of governing''.
\end{itemize}
\setstretch{2}

@@

\section*{Results}
\subsection*{Cross-Sectional OLS}

The first stage of analysis, mirroring Martin, is fitting a cross-sectional OLS regression to the data.\footnote{An inconvenient feature of both countries is a large number of ``nuisance'' candidates: fringe candidates running as independents or on behalf of minor or protest parties that have never won election, earning a tiny fraction of the vote, generally not campaigning, often non-resident or not physically present in their district. Candidates for Canada's ``Rhinoceros Party'' and New Zealand's ``McGillicuddy Serious Party'' are among those in this category. These candidates share no common support with the population of interest, and so are excluded. Because excluding these cases could be seen as ``selection on the dependent variable''---albeit in a conservative direction---I run every model presented in the main paper on the full dataset. Results on covariates of interest are not materially affected by excluding these cases.} I regress \textit{VotePct} against \textit{CabinetNow}, \textit{CabinetImportant}, and \textit{CabinetPM}, \textit{PartyInGovt}, \textit{TermsServed} and \textit{Party} fixed effects.

@@

To ensure that the association observed is not a artifact of model specifics, I test alternative specifications to ensure robustness. It is probable that a cluster structure exists in the data at the riding level, so I re-run the model with clustered standard errors at the district level and find a slightly wider but substantively similar confidence interval. The effect is also similar when substituting simple incumbency for \textit{TermsServed}, when adding district fixed effects, and when adding \textit{Age} as a proxy for candidate quality.\footnote{Coefficient estimates from the age model should generally be regarded as unreliable. Age data is available only for those who have been elected at least once, and is missing at non-random within this group; more recent members and longer-tenured members are more likely to have complete age data.} Finally, to ensure that effects are not limited to particular time periods, I disaggregate the data by decade; although standard error inflation due to lowered power makes interpretation more perilous, 6 out of 7 decades retain positive effect size estimates.

The effect's appearance in Canada contrasts its absence in New Zealand. Effect estimates in the main model are small (1.3\%) with confidence intervals spanning substantively unimportant territory. Martin's model reveals a small effect of 2.1\%, but is likely picking up correlation between tenure and cabinet selection due to the excluded tenure variable. Additional specifications are inconsistent as to whether the effect size is substantively interesting. Disaggregating by decade or election above demonstrates no consistency. High goodness of fit in relatively parsimonious models indicates that much of the variation in performance is driven by the covariates modeled (particularly party), despite limited cabinet effects. Further analysis as to why New Zealand may behave differently than Ireland or Canada is left until later.

@@

Cabinet ministers are more popular, but not necessarily because they are cabinet ministers. In fact, there is evidence to the contrary: A cross-sectional logistic regression of probability being selected into the Canadian Cabinet based on age, tenure, and previous performance suggests that previous performance matters a great deal. In the modal case, a 50 year old sophomore MP of the governing party is much 50\% more likely to be appointed a cabinet minister if she comes from a 70\% blowout win than a 40\% plurality win.

Multiple strategies exist for modelling past performance. Martin's model uses the $\Delta VotePct$ approach, where the dependent variable is the difference in vote share between elections, which assumes the following model:

@@

\setstretch{1}
\begin{eqnarray*}
	y_{i,t} = \beta X_i + \gamma y_{i,t-1} + \epsilon
\end{eqnarray*}
\newpage
\newgeometry{left=1cm,bottom=1.5cm}
\thispagestyle{fancynohead}
\begin{landscape}
\input{"Includes/model2"}
\end{landscape}
\restoregeometry

>>>>

\section*{Introduction}
Cabinet formation has long been considered from the perspective of parties and coalitions. When forming a cabinet, parties must decide who to include to please constituencies, geographies, factions within the party, and coalition partners. But parties are not monoliths, they are aggregations of members with individual preferences and goals. (M{\"u}ller and Str{\o}m 2000) Among the goals sought by any legislator are ability to achieve policy outcomes; personal and career satisfaction; service to party; and most importantly re-election. (Fenno 1978) Joining cabinet plausibly assists with the first three of these goals, but \textit{do cabinet ministers receive an electoral return in exchange for joining cabinet}? The answer is relevant to MPs deciding joining a cabinet, parties considering electoral strategy, and researchers seeking to understand the behaviour of legislators and the incentives they work with.

Past evidence suggests that governing parties incur a ``cost of governing'', which individual members can mitigate or overcome by means of developing a personal vote. (Narud and Valen 2008; Akkerman and Lange 2012) In a parliamentary context, cabinet ministers are uniquely suited to do so. This paper extends Martin (2016), who found a strong electoral effect associated with joining cabinet in Ireland. I test two national cases: Canada and New Zealand. The Canadian case reveals a robust but modest effect across specifications, while New Zealand demonstrates no consistent effect.

The paper continues by presenting past work; explaining case selection, dataset construction, and research design; presenting empirical results; and concludes by discussing possible reasons for dissonance between the Canadian and New Zealand results.

@@

Downs (1957), Key (1964), and Mayhew (1974) are among early authors who characterized legislative and party behaviour as being animated by desire for re-election. Mayhew offers three mechanisms through which legislators earn re-election: advertising, credit claiming, and position taking. Fenno (1978) takes a broader view of legislator behaviour and expands on the ways legislators form relationships with their constituents to secure re-election. An essential tension exists: on one hand, incumbent governments pay an electoral cost for governing. (Mershon 1996) On the other hand, a widely documented incumbent advantage makes staying elected an easier proposition than winning an open election. Incumbents deter competition and enjoy resource advantages as compared to their competition. (Erikson 1971; Carson 2005; Ban, Llaudet, and Snyder 2016; Gelman and King 1990)

Legislators seek to maximize their incumbency advantage while minimizing their costs of governing. Cain, Ferejohn, and Fiorina (1984) explain that constituency service can be used to build a ``personal vote''. The incentives to earn a personal vote vary by electoral and institutional context. Electoral systems such as single-member plurality where voters choose candidates directly offer the highest incentive for a personal vote, because members can be rewarded or protected by developing a reputation beyond that of their party. (Carey and Shugart 1995) The tools available to members also vary by institutional context. Parliamentary systems impose whipped voting and use vote scheduling rules to prohibit backbenchers from advancing substantive business, theoretically limiting the ability of members to distinguish themselves from the party. This has ramifications for the way in which legislators see their roles, what they hope to accomplish in a legislature, the decision to continue serving, and how they situate themselves with respect to their voters.

Studies of particularistic legislative spending find that legislators gain electoral advantage through the delivery of spending and targeted policy goods. In a US congressional context, research focuses on the ability of individual legislators to secure spending for their home districts. Parliamentary systems limit the ability of backbenchers to achieve these benefits on their own. Instead, parties act as a cartel to protect their members. In one telling of this story, the governing party uses distribution of particularistic benefits to benefit all its members. But in Martin's telling of the story, cabinet ministers have control over the shape of policy and spending within their particular ministry and use this control in concert as a cartel to support each other through the targeted provision of these goods: what he terms ``executive particularism.'' (Martin 2016) Preliminary work in a variety of national settings finds that such particularistic spending does occur at the cabinet level. (John, Ward, and Dowding 2004; Denemark 2000)

Martin's ``Policy, Office, and Votes: The Electoral Value of Ministerial Office'' is the only published work that focuses on quantifying the electoral advantage conferred by cabinet membership. Martin analyzes Irish data beginning in 1980. The empirical component tests two simple models. A cross-sectional association finds that cabinet ministers on average earn 8.5\% more of their electoral quota than like non-ministers. A second model with change in vote share as the dependent variable finds that while government members face an 8\% cost of governing, cabinet ministers erase this cost with an 8.7\% return from cabinet membership. Martin ends with a claim that ``the broad theory... is likely to be unassailable'' and an invitation to extend the analysis to different institutional and electoral environments, including mixed-member proportional systems.

In addition to executive particularism, exposure may also be a driver of electoral success. Additional media exposure and the opportunity to hold events in concert with spending or legislation may assist cabinet ministers in credit claiming. Evidence indicates that credit claiming is an important part of the ability to capitalize electorally on the delivery of spending or policy goods. (Grimmer, Messing, and Westwood 2012; Bickers, Evans, Stein, and Wrinkle 2007) To the extent that ministers enjoy a differential ability to claim credit, they should reap rewards for doing so.

\section*{Case Selection}
The ideal cases to test for the cabinet effect are Westminster parliamentary democracies with competitive multi-party elections, incumbents elected primarily from district-level elections, a norm against district-switching by incumbents, and a tendency towards single-party government. Selecting for the latter allows the research to avoid engaging the complexities of cabinet appointment in a coalition setting and intra-election coalition changes. (Laver and Shepsle 1996) I narrow my choice of cases to Australia, Canada, Ireland (studied by Martin), New Zealand, and the United Kingdom and selected Canada and New Zealand from these initial cases based on scope limitations.

Canada is a Westminster parliamentary democracy established in 1867. It features a 338 member lower house, an appointed upper house that as a matter of convention does not impede government business, and an SMP electoral system. Canada has never had a coalition government, preferring minority governments by convention. Canada has historically been governed by one of two parties: the Liberal Party and the Progressive Conservative Party (after 2004, the Conservative Party). Minor regional parties have achieved some electoral success. The effective number of political parties\footnote{A measure of party competition, where $N = \frac{1}{\sum_{i=1}^n p_i^2}$, with parties 1 through \textit{n} each occupying a $p_i$ share of seats. The measure was developed by Laakso and Taagepara (1979).} during the time studied varies from 1.5 to 3.2. Once candidates have contested an election, they do not switch ridings in subsequent elections, although parachuting (carpet-bagging) of star candidates into politically opportune ridings occurs among first-time candidates. (Koop and Bittner 2011) Elections in Canada are relatively low cost with bans on corporate and union donations, extensive public funding, and spending caps. Major party candidates typically spend in the order of \$75,000 per four-year election.

Canadian Prime Ministers enjoy broad latitude to make cabinet appointments, subject only to norms with respect to the representation of each province in cabinet, the allocation of certain portfolios to relevant geographic representatives (for example, the Fisheries ministry is traditionally given to an MP from one of Canada's coastal provinces), and increasingly diversity of gender and ethnicity. A survival analysis of cabinet entry in Canada found that gender, legal experience, education, age, and past ministerial tenure were major predictors of selection. (Kerby 2009) Over time, cabinets in Canada have become increasingly professionalized and technocratic, a trend also observed in other developed democracies (Lammers and Nyomarkay 1982; Pekkanen, Nyblade, and Krauss 2013; Berlinski, Dewan, Dowding, and Subrahmanyam 2009)

New Zealand is a Westminster parliamentary democracy established in 1852. Historically it featured a lower house with fewer than 100 members, a perfunctory upper house abolished in 1950, and an SMP electoral system with reserved electorates for voters who self-identified as indigenous Maori. Dissatisfaction with the two major political parties, National and Labour, in part because of repeated majority governments elected by weak pluralities of popular vote support, led New Zealand to enact electoral reform beginning with the 1996 national election. The current electoral system is closed-list mixed-member proportional with a low minimum vote quotas, and consists of approximately 121 seats including 71 electorate seats. The effective number of political parties varied from 1.76 to 2.16 before MMP and from 2.78 to 3.76 after MMP. Each government post-MMP has featured either supply agreements or coalitions with minor parties. Elections in New Zealand are extremely low-expense, with major party candidates spending in the order of \$20,000 per three-year election.

In principle, Labour governments in New Zealand allow their entire caucus to vote on cabinet membership, but in practice Prime Ministers are allowed considerable leeway to expand or shrink the size of cabinet and allocate portfolios among those selected. (Alley 1989; Wood 1989) Cabinets typically include representation from both of New Zealand's islands and from urban and rural areas. The Minister of Maori Affairs position is often filled by a Maori legislator, and increasingly the representation of women in cabinet has been a relevant factor. (Curtin 2015)

Both countries share a variety of cultural and institutional factors: both feature high levels of parliamentary discipline and practice cabinet collective and individual responsibility. Expulsion from cabinet, caucus, and denial of renomination are the primary punishments available to enforce party discipline. Neither country has a formal primary election mechanism, instead delegating candidate nominations to district-level parties, who by custom do not remove incumbents for reasons other than non-compliance with party leadership or scandal. (Levine 1979; Wood 1989) Finally, both cases are known to practice ``government by cabinet'', with inert backbenches and powerful cabinets. (Henderson 1989)

New Zealand and Canada differ in terms of incumbency. Canada by some measures has the lowest level of incumbency among developed countries, while New Zealand has above-average incumbency. (Matland and Studlar 2004) Incumbency has a profound effect on cabinet selection: in the Canadian low-incumbency situation, ascendant governments are forced to assemble a cabinet from caucuses comprised mostly of new, unexperienced legislators. (Kerby 2015)

\section*{Data}
The dataset I make use of is novel. Existing election datasets suffer from limitations that render them inappropriate for use here: Much of the data is aggregated above the constituency level; constituency level data is often missing names and characteristics of individual candidates necessary to make the data a panel; districts and candidates names vary across elections; and data typically does not include cabinet membership covariates.

As a result, I assembled a dataset containing full, constituency-level electoral data for all candidates and legislators in Canada from 1867 forward, and New Zealand from 1945 forward.\footnote{Canadian cabinet ministers before 1931 were expected to immediately resign from the House of Commons upon being appointed and then run in by-elections to regain the seat they resigned. To avoid the complexity of this feature of the data and the tail end of a party system transition occurring in the 1930s, I discard data before the 1945 federal election.} Canadian data was obtained by writing a series of web scrapers and harvested from Canadian Library of Parliament sources. New Zealand data was assembled using a combination of previously published volumes\footnote{Wood 1996; Norton 1988}, paper returns supplied by the NZ Electoral Commission and digitized for the first time for this project, and web scrapers for data from 1996 onwards.

@@

The paper's hypothesis is simple: Cabinet ministers enjoy an electoral advantage compared to backbenchers.

The primary limitation when testing this hypothesis is severe endogeneity. The ``gold standard'' to tackle this problem would be to experimentally appoint cabinet ministers through random assignment and measure causal effects. Reality is not so kind: Parties select cabinet ministers to satisfy goals including rewarding key constituencies, demographics, and policy demanders within the party and coalition partners. Moreover, parties have access to private signals about incumbent quality learned during the recruitment, vetting, and campaigning process that are not fully accounted for in vote share at the time of their first election. These latent candidate qualities may become obvious during the process of legislating and materialize in future results, but are not strictly speaking a cabinet effect: Koop and Bittner (2011) argue that ``star candidates'' are much more likely to be appointed to cabinet. In light of these concerns, the claims made here cannot be said to be causal: effects in this paper are associations, suggestive of a relationship, but not decisive.

Tables in this paper present a main model and a model that ``replicates'' Martin's comparable design. I run alternative specifications to test robustness of the observed effects and include these in Appendices 1-3.

@@

George Box remarked that ``All models are wrong, but some are useful''. This aphorism is a reminder that no model can capture all sources of variation, but even misspecified models can uncover useful patterns in the data at hand to test theory. In keeping with this observation, it is likely that the models presented here omit variables due to limitations of data, theory, or imagination. Of particular note, the scope of this project rendered it impossible to collect richer biographical covariate data on candidates. Subsequent research that can gather such data could explain residual variation in candidate quality and more clearly isolate the effect of interest or allow a more robust selection on observables causal identification strategy. In the mean time, the modeled variables are as follows:

\setstretch{1.5}
\begin{itemize}
\item \textit{VotePct}: The vote share of the candidate.
\item \textit{CabinetNow}: The candidate was a member of the cabinet during the preceding term. I exclude ministers outside cabinet, junior ministers, ministers without portfolio, and other quasi-cabinet roles but include deputy Prime Ministers and Attorneys General.\footnote{I also do not model ministerial exit from cabinet or role-switching mid-term, so in principle some of the \textit{CabinetNow} incumbents have been removed from cabinet before the election. Evidence suggests that ministerial exit is predicted by poor performance (Berlinkski, Dewan, and Dowding 2010) or scandal, (Dewan and Myatt 2007; Dewan and Dowding 2005) so this modeling decision should cause effect estimates to be biased conservatively as poor ministers drag the overall effect down.}
\item \textit{CabinetImportant}: The candidate was an ``important'' minister of the cabinet during the preceding term. ``Importance'' is subjective and here includes Defence, Health, Justice, Finance, Public Works, Revenue, and Foreign ministries. This variable exists to capture heterogeneity in the cabinet effect; senior ministers are presumed to have more discretion with respect to particularistic spending, more control over policy outcomes, and greater capacity to cultivate a personal vote through media exposure and credit claiming.
\item \textit{CabinetPM}: Dummy variable for the Prime Minister, regardless of whether they were an incumbent member of the legislature. The inclusion of a control for Prime Ministers is intended to absorb the positive electoral effects expected for a sitting Prime Minister to avoid their misattribution to cabinet membership.
\item \textit{Incumbent}: The candidate is an incumbent member of the legislature. I treat candidates as incumbent even if redistricting occurs, or if the candidate was previously elected by party list and now contests an electorate. Incumbency is known to be a a strong predictor of re-election.
\item \textit{TermsServed}: The total unbroken number of terms served by a candidate. Tenure does not reset if an incumbent resigns and runs in a by-election, but does reset if a candidate loses or resigns and later returns to the legislature. I treat by-elected members as serving a full time. Tenure is present in the model to capture the accumulated effects of developing a personal vote (or, at least, a reputation of competency) in an incumbent's district over time.
\item \textit{Age}: The age of the candidate at the time of the election. Only available for a subset of Canadian candidates who were elected to office at least once. Age was gathered to allow for a proxy for signals of candidate quality including occupation, past electoral and career experience.
\item \textit{Elected}: Binary indicator for election outcome.
\item \textit{Party}: Party label the member stood for election under. I recode minor parties or splinter factions to be independent.
\item \textit{PartyInGovt}: Is the candidate running for the party currently governing? In New Zealand post-1996, I include junior coalition partner parties but not parties in supply agreements with the governing party. This variable was constructed in order to account for the ``electoral costs of governing''.
\end{itemize}
\setstretch{2}

@@

\section*{Results: Cross-Sectional OLS}

The first stage of analysis, mirroring Martin, is fitting a cross-sectional OLS regression to the data.\footnote{An inconvenient feature of both countries is a large number of ``nuisance'' candidates: fringe candidates running as independents or on behalf of minor or protest parties that have never won election, earning a tiny fraction of the vote, generally not campaigning, often non-resident or not physically present in their district. Candidates for Canada's ``Rhinoceros Party'' and New Zealand's ``McGillicuddy Serious Party'' are among those in this category. These candidates share no common support with the population of interest, and so are excluded. Because excluding these cases could be seen as ``selection on the dependent variable''---albeit with no common support on the coefficient of interest---I run every model presented in the main paper on the full dataset. Results on covariates of interest are not materially affected by excluding these cases.} I regress \textit{VotePct} against \textit{CabinetNow}, \textit{CabinetImportant}, and \textit{CabinetPM}, \textit{PartyInGovt}, \textit{TermsServed} and \textit{Party} fixed effects.

@@

To ensure that the association observed is not a artifact of model specifics, I test alternative specifications to ensure robustness. It is probable that a cluster structure exists in the data at the riding level, so I re-run the model with clustered standard errors at the district level and find a slightly wider but substantively similar confidence interval. The effect is also similar when substituting simple incumbency for \textit{TermsServed}, when adding district fixed effects, and when adding \textit{Age} as a proxy for candidate quality.\footnote{Coefficient estimates from the age model should generally be regarded as unreliable. Age data is available only for those who have been elected at least once, and is missing at non-random within this group; more recent members and longer-tenured members are more likely to have complete age data.} Finally, to ensure that effects are not limited to particular time periods, I disaggregate the data by decade; although lowered power makes interpretation more perilous, 6 out of 7 decades retain positive effect size estimates.

The effect's appearance in Canada contrasts its absence in New Zealand. Effect estimates in the main model are small (1.3\%) with confidence intervals spanning substantively unimportant territory. Martin's model reveals a small effect of 2.1\%, but is likely picking up correlation between tenure and cabinet selection due to the excluded tenure variable. Additional specifications are inconsistent as to whether the effect size is substantively interesting. Disaggregating by decade reveals no consistency. The relatively parsimonious models have high goodness of fit, indicating that much of the variation in performance is driven by the covariates modeled, particularly party label, despite limited cabinet effects. Further analysis as to why New Zealand may behave differently than Ireland or Canada is left until after presentation of results.

@@

Cabinet ministers are more popular, but not necessarily because they are cabinet ministers. In fact, there is evidence the causality is exactly reversed: A cross-sectional logistic regression of probability being selected into the Canadian Cabinet based on age, tenure, and previous performance suggests that previous performance matters a great deal. In the modal case, a 50 year old sophomore MP of the governing party is much 50\% more likely to be appointed a cabinet minister if she comes from a 70\% blowout win than a 40\% plurality win.

Multiple strategies exist for modeling past performance. Martin's model uses the $\Delta VotePct$ approach, where the dependent variable is the difference in vote share between elections, which assumes the following model:

@@

\setstretch{1}
\begin{eqnarray*}
	y_{i,t} = \beta X_i + \gamma y_{i,t-1} + \epsilon
\end{eqnarray*}

\newpage
\newgeometry{left=1cm,bottom=1.5cm}
\thispagestyle{fancynohead}
\begin{landscape}
\input{"Includes/model2"}
\end{landscape}
\restoregeometry
